\pkg{cv}: An \proglang{R} Package for Cross-Validation of Regression Models

### 2024-01-11 scratch/notes_john/JSS-article/JSS-article-reduced.Rmd
small edits to JSS paper

```{r mpg-horsepower-scatterplot-polynomials, echo=FALSE}
#| out.width = "50%",
#| fig.height = 5,
#| fig.cap = "mpg vs horsepower for the Auto data, showing fitted polynomials of degree 1 through 5."
plot(mpg ~ horsepower, data=Auto)
horsepower <- with(Auto, 
                   seq(min(horsepower), max(horsepower), 
                       length=1000))
for (p in 1:5){
  m <- lm(mpg ~ poly(horsepower,p), data=Auto)
  mpg <- predict(m, newdata=data.frame(horsepower=horsepower))
  lines(horsepower, mpg, col=p + 1, lty=p, lwd=2)
}
legend("topright", legend=1:5, col=2:6, lty=1:5, lwd=2,
       title="Degree", inset=0.02)
```

@@

# LOO CV
cv.auto.loo <- cv(mlist, data=Auto, k="loo")
cv.auto.loo[2] # e.g., for quadratic model
```
The \code{models()} function takes an arbitrary number of regression models as its arguments, which are optionally named, to create a \code{"modList"} object. Because we generated the polynomial regression models in a named list, we conveniently employ \code{do.call()} to supply the models as arguments to \code{models()}. Note that the names created for the list (e.g., \code{"m.2"}) are then used for the models. We can also invoke the \code{plot()} method for \code{"cvModList"} objects to compare the models (see Figure \ref{fig:polynomial-regression-CV-graph-2}):

@@

Some data-management is required before fitting a mixed-effects model to the HSB data:
```{r HSB-data-management, cache=TRUE}
HSB <- MathAchieve
HSB <- merge(MathAchSchool[, c("School", "Sector")],
             HSB[, c("School", "SES", "MathAch")], by="School")
names(HSB) <- tolower(names(HSB))
HSB <- within(HSB, {
  mean.ses <- ave(ses, school)
  cses <- ses - mean.ses
})
```
In the process, we merged variables from the school-level and student-level data sets and created two new school-level variables: \code{mean.ses}, which is the average SES for students in each school; and \code{cses}, which is each students' SES centered at their school means. For details, see @FoxWeisberg:2019 [Sec. 7.2.2].

>>>>

```{r mpg-horsepower-scatterplot-polynomials, echo=FALSE}
#| out.width = "50%",
#| fig.height = 5,
#| fig.cap = "mpg vs horsepower for the Auto data, showing fitted polynomials of degree 1 through 5."
plot(mpg ~ horsepower, data=Auto)
horsepower <- with(Auto, 
                   seq(min(horsepower), max(horsepower), 
                       length=1000))
for (p in 1:5){
  m <- lm(mpg ~ poly(horsepower,p), data=Auto)
  mpg <- predict(m, newdata=data.frame(horsepower=horsepower))
  lines(horsepower, mpg, col=p + 1, lty=p, lwd=4)
}
legend("topright", legend=1:5, col=2:6, lty=1:5, lwd=4,
       title="Degree", inset=0.02)
```

@@

# LOO CV
cv.auto.loo <- cv(mlist, data=Auto, k="loo")
cv.auto.loo[2] # e.g., for quadratic model
```
The \code{models()} function takes an arbitrary number of regression models as its arguments, which are optionally named, to create a \code{"modList"} object. Because we generated the polynomial regression models in a named list, we conveniently employ \code{do.call()} to supply the models as arguments to \code{models()}. The names created for the list (e.g., \code{"m.2"}) are then used for the models. We can also invoke the \code{plot()} method for \code{"cvModList"} objects to compare the models (see Figure \ref{fig:polynomial-regression-CV-graph-2}):

@@

Some data-management is required before fitting a mixed-effects model to the HSB data:
```{r HSB-data-management, cache=TRUE}
HSB <- MathAchieve
HSB <- merge(MathAchSchool[, c("School", "Sector")],
             HSB[, c("School", "SES", "MathAch")], by="School")
names(HSB) <- tolower(names(HSB))
HSB <- within(HSB, {
  mean.ses <- ave(ses, school)
  cses <- ses - mean.ses
})
```
In the process, we merged variables from the school-level and student-level data sets and created two new school-level variables: \code{mean.ses}, which is the average SES for students in each school; and \code{cses}, which is the individual students' SES centered at their school means. For details, see @FoxWeisberg:2019 [Sec. 7.2.2].

### 2024-01-11 scratch/notes_john/JSS-article/JSS-article-reduced.tex
small edits to JSS paper

Coefficients:
                      Estimate Std. Error t value Pr(>|t|)    
(Intercept)            23.4459     0.2209  106.13   <2e-16 ***
poly(horsepower, 2)1 -120.1377     4.3739  -27.47   <2e-16 ***
poly(horsepower, 2)2   44.0895     4.3739   10.08   <2e-16 ***
---
Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1

Residual standard error: 4.374 on 389 degrees of freedom
Multiple R-squared:  0.6876,    Adjusted R-squared:  0.686 
F-statistic:   428 on 2 and 389 DF,  p-value: < 2.2e-16
\end{CodeOutput}
\begin{CodeInput}
R> cv(m.auto, confint=TRUE)
\end{CodeInput}
\begin{CodeOutput}
R RNG seed set to 542508
\end{CodeOutput}
\begin{CodeOutput}
10-Fold Cross Validation
method: Woodbury
criterion: mse
cross-validation criterion = 19.57393
bias-adjusted cross-validation criterion = 19.54281
95% CI for bias-adjusted CV criterion = (16.03528, 23.05034)
full-sample criterion = 18.98477 
\end{CodeOutput}
\end{CodeChunk}

@@

The \code{models()} function takes an arbitrary number of regression
models as its arguments, which are optionally named, to create a
\code{"modList"} object. Because we generated the polynomial regression
models in a named list, we conveniently employ \code{do.call()} to
supply the models as arguments to \code{models()}. Note that the names
created for the list (e.g., \code{"m.2"}) are then used for the models.
We can also invoke the \code{plot()} method for \code{"cvModList"}
objects to compare the models (see Figure
\ref{fig:polynomial-regression-CV-graph-2}):

@@

In the process, we merged variables from the school-level and
student-level data sets and created two new school-level variables:
\code{mean.ses}, which is the average SES for students in each school;
and \code{cses}, which is each students' SES centered at their school
means. For details, see \citet[Sec. 7.2.2]{FoxWeisberg:2019}.

@@

Coefficients:
            Estimate Std. Error t value Pr(>|t|)    
(Intercept) 9.050430   0.347189  26.068  < 2e-16 ***
x           0.020908   0.007273   2.875  0.00422 ** 
---
Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1

@@

Coefficients:
            Estimate Std. Error t value Pr(>|t|)    
(Intercept)  9.93704    0.03122   318.3   <2e-16 ***
---
Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1

@@

Coefficients:
            Estimate Std. Error t value Pr(>|t|)    
(Intercept)  9.93716    0.03098 320.803  < 2e-16 ***
X.99        -0.09103    0.03082  -2.953  0.00322 ** 
X.90        -0.08205    0.03135  -2.617  0.00901 ** 
X.87        -0.06942    0.03105  -2.235  0.02561 *  
X.40        -0.04759    0.03076  -1.547  0.12211    
X.65        -0.05523    0.03147  -1.755  0.07952 .  
X.91         0.05245    0.03084   1.700  0.08937 .  
X.53        -0.04921    0.03048  -1.615  0.10672    
X.45         0.05543    0.03182   1.742  0.08183 .  
X.31         0.04525    0.03108   1.456  0.14570    
X.56         0.05433    0.03273   1.660  0.09723 .  
X.61        -0.05085    0.03170  -1.604  0.10908    
X.60        -0.05133    0.03194  -1.607  0.10832    
X.46         0.05158    0.03272   1.576  0.11529    
X.35         0.04696    0.03146   1.493  0.13584    
X.92         0.04430    0.03100   1.429  0.15329    
---
Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1

@@

Coefficients:
              Estimate Std. Error t value Pr(>|t|)    
(Intercept)   9.434594   0.261529  36.075  < 2e-16 ***
horsepower   -0.276254   0.056143  -4.921 1.30e-06 ***
weight       -0.609071   0.056003 -10.876  < 2e-16 ***
acceleration -0.131380   0.053195  -2.470  0.01397 *  
year71        0.027984   0.028936   0.967  0.33412    
year72       -0.007111   0.028446  -0.250  0.80274    
year73       -0.039529   0.026014  -1.520  0.12947    
year74        0.052752   0.029986   1.759  0.07936 .  
year75        0.053199   0.029280   1.817  0.07004 .  
year76        0.074317   0.028212   2.634  0.00878 ** 
year77        0.137931   0.028875   4.777 2.56e-06 ***
year78        0.145876   0.027529   5.299 1.99e-07 ***
year79        0.236036   0.029080   8.117 6.99e-15 ***
year80        0.335274   0.031148  10.764  < 2e-16 ***
year81        0.262872   0.030555   8.603  < 2e-16 ***
year82        0.323391   0.029608  10.922  < 2e-16 ***
originEurope  0.055818   0.016785   3.326  0.00097 ***
originJapan   0.043554   0.017479   2.492  0.01314 *  
---
Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1

@@

\begin{CodeChunk}
\begin{CodeInput}
R> BayesRule
\end{CodeInput}
\begin{CodeOutput}
function(y, yhat){
  if (!all(y %in% c(0, 1))) stop("response values not all 0 or 1")
  if (any(yhat < 0) || any(yhat > 1)) stop("fitted values outside of interval [0, 1]")
  yhat <- round(yhat)
  result <- mean(y != yhat) # proportion in error
  attr(result, "casewise loss") <- "y != round(yhat)"
  result
}
<bytecode: 0x129d04270>
<environment: namespace:cv>
\end{CodeOutput}
\end{CodeChunk}

>>>>

Coefficients:
                      Estimate Std. Error t value Pr(>|t|)
(Intercept)            23.4459     0.2209  106.13   <2e-16
poly(horsepower, 2)1 -120.1377     4.3739  -27.47   <2e-16
poly(horsepower, 2)2   44.0895     4.3739   10.08   <2e-16

Residual standard error: 4.374 on 389 degrees of freedom
Multiple R-squared:  0.6876,    Adjusted R-squared:  0.686 
F-statistic:   428 on 2 and 389 DF,  p-value: < 2.2e-16
\end{CodeOutput}
\begin{CodeInput}
R> cv(m.auto, confint=TRUE)
\end{CodeInput}
\begin{CodeOutput}
R RNG seed set to 32925
\end{CodeOutput}
\begin{CodeOutput}
10-Fold Cross Validation
method: Woodbury
criterion: mse
cross-validation criterion = 19.41501
bias-adjusted cross-validation criterion = 19.3924
95% CI for bias-adjusted CV criterion = (15.88674, 22.89806)
full-sample criterion = 18.98477 
\end{CodeOutput}
\end{CodeChunk}

@@

The \code{models()} function takes an arbitrary number of regression
models as its arguments, which are optionally named, to create a
\code{"modList"} object. Because we generated the polynomial regression
models in a named list, we conveniently employ \code{do.call()} to
supply the models as arguments to \code{models()}. The names created for
the list (e.g., \code{"m.2"}) are then used for the models. We can also
invoke the \code{plot()} method for \code{"cvModList"} objects to
compare the models (see Figure
\ref{fig:polynomial-regression-CV-graph-2}):

@@

In the process, we merged variables from the school-level and
student-level data sets and created two new school-level variables:
\code{mean.ses}, which is the average SES for students in each school;
and \code{cses}, which is the individual students' SES centered at their
school means. For details, see \citet[Sec. 7.2.2]{FoxWeisberg:2019}.

@@

Coefficients:
            Estimate Std. Error t value Pr(>|t|)
(Intercept) 9.050430   0.347189  26.068  < 2e-16
x           0.020908   0.007273   2.875  0.00422

@@

Coefficients:
            Estimate Std. Error t value Pr(>|t|)
(Intercept)  9.93704    0.03122   318.3   <2e-16

@@

Coefficients:
            Estimate Std. Error t value Pr(>|t|)
(Intercept)  9.93716    0.03098 320.803  < 2e-16
X.99        -0.09103    0.03082  -2.953  0.00322
X.90        -0.08205    0.03135  -2.617  0.00901
X.87        -0.06942    0.03105  -2.235  0.02561
X.40        -0.04759    0.03076  -1.547  0.12211
X.65        -0.05523    0.03147  -1.755  0.07952
X.91         0.05245    0.03084   1.700  0.08937
X.53        -0.04921    0.03048  -1.615  0.10672
X.45         0.05543    0.03182   1.742  0.08183
X.31         0.04525    0.03108   1.456  0.14570
X.56         0.05433    0.03273   1.660  0.09723
X.61        -0.05085    0.03170  -1.604  0.10908
X.60        -0.05133    0.03194  -1.607  0.10832
X.46         0.05158    0.03272   1.576  0.11529
X.35         0.04696    0.03146   1.493  0.13584
X.92         0.04430    0.03100   1.429  0.15329

@@

Coefficients:
              Estimate Std. Error t value Pr(>|t|)
(Intercept)   9.434594   0.261529  36.075  < 2e-16
horsepower   -0.276254   0.056143  -4.921 1.30e-06
weight       -0.609071   0.056003 -10.876  < 2e-16
acceleration -0.131380   0.053195  -2.470  0.01397
year71        0.027984   0.028936   0.967  0.33412
year72       -0.007111   0.028446  -0.250  0.80274
year73       -0.039529   0.026014  -1.520  0.12947
year74        0.052752   0.029986   1.759  0.07936
year75        0.053199   0.029280   1.817  0.07004
year76        0.074317   0.028212   2.634  0.00878
year77        0.137931   0.028875   4.777 2.56e-06
year78        0.145876   0.027529   5.299 1.99e-07
year79        0.236036   0.029080   8.117 6.99e-15
year80        0.335274   0.031148  10.764  < 2e-16
year81        0.262872   0.030555   8.603  < 2e-16
year82        0.323391   0.029608  10.922  < 2e-16
originEurope  0.055818   0.016785   3.326  0.00097
originJapan   0.043554   0.017479   2.492  0.01314

@@

\begin{CodeChunk}
\begin{CodeInput}
R> BayesRule
\end{CodeInput}
\begin{CodeOutput}
function(y, yhat){
  if (!all(y %in% c(0, 1))) stop("response values not all 0 or 1")
  if (any(yhat < 0) || any(yhat > 1)) stop("fitted values outside of interval [0, 1]")
  yhat <- round(yhat)
  result <- mean(y != yhat) # proportion in error
  attr(result, "casewise loss") <- "y != round(yhat)"
  result
}
<bytecode: 0x13e7ed098>
<environment: namespace:cv>
\end{CodeOutput}
\end{CodeChunk}
